Prediction of PIK3CA mutation with gene expression

### 2020-04-03 plosOne/plosOne.Rmd
discussion

```{r}
knitr::read_chunk('../data.R')
knitr::read_chunk('../prediction.R')
knitr::read_chunk('../logisticRegression.R')
```

@@

RNAseq can be widely used. The mutation status directly predictio

@@

The proposed solution

@@

PIK3CA mutation data was get using cgdsr rpackage. Gene expression data was get from GDAC firehose using RTCGAToolbox R package. ER immunostain postive HER2 immunostain negative and/or SISH negative breast cancers are included. Data of invasive ductal carcinomas were used for training set and data of invasive lobular carcinoma were used for test set. Number of observations were 530 in training set and 188 in test set.

@@

To narrow down potential predictors, Genes with a large variance (more than median) and differently expressed by PIK3CA mutation status (wilcox test, *P* $<10^{-8}$) were selected. 111 out of 20502 genes were included in the modeling process.

@@

BoxCox transformation was done to correct skewness. Centering and scaling were done. All preprocessing was done using recipe r package.

@@

```{r load}
<<load>>
```

```{r}
<<modeling>>
```

```{r}
<<Finalize_model>>

```

Table 1. Frequency of PIK3CA mutation in training set and test set

```{r}
library(tidyverse)
train_target <- tibble::tibble(summary(trainSetDiff$PIK3CA_T)) 
test_target <- tibble::tibble(summary(testSetDiff$PIK3CA_T))
target <- bind_cols(train_target, test_target) %>%
  rename(TrainSet = "summary(trainSetDiff$PIK3CA_T)", TestSet = "summary(testSetDiff$PIK3CA_T)")
knitr::kable(target)

```

```{r}
<<trainset_prediction>>

knitr::kable(confusion_matrix$table)
```

## selecting model
Cross-validation 
The model showed best performance at lambda = 0.01 and alpha = 1.0 (Ridge regression).

@@

@@

Critical assessment opinions on 
- any shortcomings in study design  
- limitations in methods
- flaws in analysis
- validity of assumption

@@

>>>>

```{r}
knitr::read_chunk('../prediction.R')
```

@@

RNAseq can be widely used. The mutation status directly prediction

@@

prediction of aberant activation of a certain pathway vs a specific mutation
The proposed solution

Prediction RAS pathway activation with gene expression data was done in previous study. @way_2018_machine_cellreports They trained pancancer The Cancer Genome Atlas (TCGA) data with a supervised elastic net penalized logistic regression classifier with stochastic gradient descent. The performance of their model was 84% with an area under the receiver operating characteristic (AUROC) curve and 63% with an area under the precision recall (AUPR) curve. We applied their modeling methods in prediction PIK3CA mutation.

@@

We used TCGA pancancer dataset. PIK3CA mutation data was get using cgdsr rpackage. Gene expression data was get from GDAC firehose using RTCGAToolbox R package. ER immunostain postive HER2 immunostain negative and/or SISH negative breast cancers are included. Data of invasive ductal carcinomas were used for training set and data of invasive lobular carcinoma were used for test set. Number of observations were 530 in training set and 188 in test set.

@@

To narrow down potential predictors, Genes with a large the median absolute deviation (more than third-quartiles) were selected. 5000 out of 20502 genes were included in the modeling process.

@@

Yeo-Johnson transformation was done to correct skewness. Centering and scaling were done. All preprocessing was done using recipe r package.

@@

## selecting model

Cross-validation  
The model showed best performance at lambda = 0.01 and alpha = 1.0 (Ridge regression).

@@

Our model showed good performance to predict PIK3CA mutation of various cancer types.

@@

Critical assessment opinions on 
- any shortcomings in study design

In this study, the split method was done for testset instead of indepedent testset. Because there is no standard analysis method for the RNA-seq gene expression quantification, it is difficult to find gene expression data same as TGCA data. Developing standard analysis method for gene expression data is neccessory to apply this prediction model.

- limitations in methods

Case imbalance

- flaws in analysis

- validity of assumption

@@

# Figure legends

### 2020-04-03 plosOne/plosOne.tex
discussion

The proposed solution

@@

PIK3CA mutation data was get using cgdsr rpackage. Gene expression data
was get from GDAC firehose using RTCGAToolbox R package. ER immunostain
postive HER2 immunostain negative and/or SISH negative breast cancers
are included. Data of invasive ductal carcinomas were used for training
set and data of invasive lobular carcinoma were used for test set.
Number of observations were 530 in training set and 188 in test set.

@@

@@

Critical assessment opinions on - any shortcomings in study design\\
- limitations in methods - flaws in analysis - validity of assumption

@@

>>>>

prediction of aberant activation of a certain pathway vs a specific
mutation The proposed solution

Prediction RAS pathway activation with gene expression data was done in
previous study. {[}2{]} They trained pancancer The Cancer Genome Atlas
(TCGA) data with a supervised elastic net penalized logistic regression
classifier with stochastic gradient descent. The performance of their
model was 84\% with an area under the receiver operating characteristic
(AUROC) curve and 63\% with an area under the precision recall (AUPR)
curve. We applied their modeling methods in prediction PIK3CA mutation.

@@

We used TCGA pancancer dataset. PIK3CA mutation data was get using cgdsr
rpackage. Gene expression data was get from GDAC firehose using
RTCGAToolbox R package. ER immunostain postive HER2 immunostain negative
and/or SISH negative breast cancers are included. Data of invasive
ductal carcinomas were used for training set and data of invasive
lobular carcinoma were used for test set. Number of observations were
530 in training set and 188 in test set.

\hypertarget{selecting-variable-for-modeling}{%
\subsection{Selecting variable for
modeling}\label{selecting-variable-for-modeling}}

To narrow down potential predictors, Genes with a large the median
absolute deviation (more than third-quartiles) were selected. 5000 out
of 20502 genes were included in the modeling process.

\hypertarget{preprocessing}{%
\subsection{Preprocessing}\label{preprocessing}}

Yeo-Johnson transformation was done to correct skewness. Centering and
scaling were done. All preprocessing was done using recipe r package.

\hypertarget{modeling}{%
\subsection{Modeling}\label{modeling}}

Penalized logistic regression was applied to prediction modeling.
10-fold cross-validation with targe variable stratification was done
over the hyperparameter grid: \(\lambda\) \{\(10^{-5}\),
\(10^{-4}\),\(10^{-3}\),\(10^{-2}\),\(10^{-1}\), \(10^{0}\)\},
\(\alpha\) \{0.0, 0.25, 0.5, 0.75\}. Model performance was evealuated
with the area under receiver operating characteristic (ROC) and area
under the ROC (AUROC) and precision recall (PR) and area under curve
(AUPR).

@@

\hypertarget{selecting-model}{%
\subsection{selecting model}\label{selecting-model}}

Cross-validation\\
The model showed best performance at lambda = 0.01 and alpha = 1.0
(Ridge regression).

\hypertarget{prediction-performance}{%
\subsection{Prediction performance}\label{prediction-performance}}

\hypertarget{performance-of-each-cancer-type}{%
\subsection{performance of each cancer
type}\label{performance-of-each-cancer-type}}

\hypertarget{important-predictors}{%
\subsection{Important predictors}\label{important-predictors}}

@@

Critical assessment opinions on - any shortcomings in study design

In this study, the split method was done for testset instead of
indepedent testset. Because there is no standard analysis method for the
RNA-seq gene expression quantification, it is difficult to find gene
expression data same as TGCA data. Developing standard analysis method
for gene expression data is neccessory to apply this prediction model.

\begin{itemize}
\tightlist
\item
  limitations in methods
\end{itemize}

Case imbalance

\begin{itemize}
\item
  flaws in analysis
\item
  validity of assumption
\end{itemize}

@@

\hypertarget{figure-legends}{%
\section*{Figure legends}\label{figure-legends}}
\addcontentsline{toc}{section}{Figure legends}
